---
title: "campaign finance"
output: 
---

FEC bulk data URL: https://www.fec.gov/data/browse-data/?tab=bulk-data
Data from the 2021-2022 election cycle

```{r}
library(tidyverse)
library(lubridate)
```

```{r}
transactions <- read_csv("data/transactions.csv")
transaction_types <- read_csv("data/transaction_types.csv")
candidates <- read_csv("data/candidates.csv")
committees <- read_csv("data/committees.csv")
```

What transaction types are we working with? 
```{r}
transactions %>% 
  count(transaction_tp)
```

Join to transaction_types
Test first
```{r}
transaction_types %>% 
  count(transaction_tp) %>% 
  filter(n > 1)

# no duplicates
```

Left join
```{r}
# first check: are the results = rows in main table? 
transactions %>% 
  left_join(transaction_types, by = "transaction_tp")

# second check: are there any rows where the joined columns are NA?
transactions %>% 
  left_join(transaction_types, by = "transaction_tp") %>% 
  filter(is.na(transaction_desc))
# no missing values

# check the code and desc matchups
transactions %>% 
  left_join(transaction_types, by = "transaction_tp") %>% 
  count(transaction_tp, transaction_desc)

# optional step is to create a new table from the join:
transactions2 <- transactions %>% 
  left_join(transaction_types, by = "transaction_tp")
```

How much money was spent for each transaction type? 
```{r}
transactions %>% 
  left_join(transaction_types, by = "transaction_tp") %>% 
  group_by(transaction_desc) %>% 
  summarise(total_amt = sum(transaction_amt), n = n()) %>% 
  arrange(desc(total_amt))
```

Which candidates are in our data? 
```{r}
transactions %>% 
  count(cand_id) %>% 
  arrange(desc(n)) %>% 
  left_join(candidates, by = "cand_id")
```

Which candidate had the most money spent on them in independent expenditures?
```{r}
transactions %>% 
  filter(transaction_tp == "24A" | transaction_tp == "24E") %>%
  left_join(candidates, by = "cand_id") %>% 
  group_by(cand_name) %>% 
  summarise(total_amt = sum(transaction_amt)) %>% 
  arrange(desc(total_amt))
```
```{r}
# look at opposing
transactions %>% 
  filter(transaction_tp == "24A") %>%
  left_join(candidates, by = "cand_id") %>% 
  group_by(cand_name) %>% 
  summarise(total_amt = sum(transaction_amt)) %>% 
  arrange(desc(total_amt))

# look at advocating
transactions %>% 
  filter(transaction_tp == "24E") %>%
  left_join(candidates, by = "cand_id") %>% 
  group_by(cand_name) %>% 
  summarise(total_amt = sum(transaction_amt)) %>% 
  arrange(desc(total_amt))
```

Can we make this code more efficient so we're not repeating ourselves?
(The DRY principle: Don't Repeat Yourself)
The function `pivot_wider()` can help us here: 
```{r}
transactions %>% 
  filter(transaction_tp %in% c("24A", "24E")) %>% 
  left_join(candidates, by = "cand_id") %>% 
  group_by(cand_name, transaction_tp) %>% 
  summarise(total_amt = sum(transaction_amt)) %>% 
  pivot_wider(names_from = transaction_tp, values_from = total_amt) %>% 
  mutate(total = sum(`24A`,`24E`, na.rm=T), opposing_pct = `24A`/total) %>% 
  arrange(desc(opposing_pct))
```

Who are the committees spending this independent expenditure money? 
```{r}
transactions %>% 
  filter(transaction_tp %in% c("24A", "24E")) %>% 
  left_join(committees, by = "cmte_id") %>% 
  group_by(cmte_nm) %>% 
  summarise(total = sum(transaction_amt)) %>% 
  arrange(desc(total))
```

