---
title: "Columbia Dispatch"
output: html_notebook
editor_options: 
  markdown: 
    wrap: 72
---

Columbia Police Dispatch:
<https://www.como.gov/CMS/911dispatch/police.php>

-   Downloaded 1/29/2024 8:43 a.m.

-   1/1/2019 to 1/29/2024 (This is what we requested, but we actually
    only got: 1/29/2023 to 1/29/2024)

```{r}
library(tidyverse)
library(lubridate)
```

Load the data

```{r}
dispatch <- read_csv("data/como_dispatch.csv")
```

Get to know your data: have a routine that you execute every time you
open up a new dataset.\
Three questions I always ask:

1.  How many rows? 83,190 rows

2.  What is one row? 1 row = 1 call that resulted in a dispatch (we
    think, though there are some duplicates we need to ask about)

3.  What is in every column...

Assessing the data using `count()`

```{r}
# Are there any incidents with the exact same date, time and address?
dispatch %>% 
  group_by(CallDateTime, Address) %>% 
  summarise(n = n()) %>% 
  filter(n > 1)
# There are 50 uniqe date-time / address combinations with more than one record.

# Another way to look at this uses the `count()` function to replace the `group_by()` and `summarise(n = n())`:
dispatch %>% 
  count(CallDateTime, Address) %>% 
  filter(n > 1)
# the results are exactly the same as above.

# Take a look at one example:
dispatch %>% 
  filter(CallDateTime == "10/8/2023 2:08:54 PM")
# these four rows all have the same info except different incident numbers. different tickets? 

# Does InNum have any duplicates? 
dispatch %>% 
  count(InNum) %>% 
  arrange(desc(n))
# Yes, take a look at one example: 

dispatch %>% 
  filter(InNum == "2023029401") 
# These appear to be exact duplicates.
```

There are exact duplicates! - call the city

Let's say we learned that exact duplicates are mistakes and should be
removed:
```{r}
# The `distinct()` function shows distinct rows:
dispatch %>% 
  distinct()
# 20 records fewer

# Remove those 20 records permanently:
dispatch <- dispatch %>% 
  distinct()
```

Assess every column:

```{r}
dispatch %>% 
  count(InNum) %>% 
  arrange(desc(n))
# Still some duplicates

dispatch %>% 
  count(ExtNatureDisplayName)
# lots of values here, some are hard to distinguish.

dispatch %>% 
  count(PolArea)
# These are useful if we find a map of them: https://www.como.gov/wp-content/uploads/2022/01/Beat-map-2.pdf

dispatch %>% 
  count(DOW)
# clean, and incidents are pretty evenly distributed (except for Sunday?)
```

We can't really assess the CallDateTime because it's a date stored as text. We must convert it to a true date; to do this cleaning, we'll create a new column with the dates using the `mutate()` function and a date conversion function from the package `lubridate`:
```{r}
# first test your conversion to make sure it worked
dispatch %>% 
  mutate(new_date = mdy_hms(CallDateTime)) %>% 
  select(CallDateTime, new_date)

# next overwrite your `dispatch` variable so it has the new column
dispatch <- dispatch %>% 
  mutate(new_date = mdy_hms(CallDateTime))
```

What's the date range of our data?

```{r}
range(dispatch$new_date)
```

Other ways to work with true dates:

```{r}
dispatch %>% 
  count(month(new_date))
```
