---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```

Using read_csv, let's talk about functions. Pull up the documentation for a function by running `?function_name`

```{r}
?read_csv
```

The documentation will show up in the bottom right panel.

There are several functions in this grouping: scroll down to `read_csv`

The arguments are listed below it. Any argument that isn't followed by a `=` is required. Everything that has a default setting is optional; only use it if you don't want to use the default.

Messing with column names:

```{r}
# make it so that R doesn't read the first row as a header row:
read_csv("data/como_dispatch.csv", col_names = FALSE)

# give R a list of headers for your data:
read_csv("data/como_dispatch.csv", 
         col_names = c("col1", "col2", "col3", "col4", "col5", "col6", "col7", "col8"))

# give R a list of headers and skip the first row (which you might do if the headers were really bad):
read_csv("data/como_dispatch.csv", 
         col_names = c("col1", "col2", "col3", "col4", "col5", "col6", "col7", "col8"),
         skip = 1)
```

Data types

```{r}
# specify a column type for each column in a string (each character represents one column, in order):
read_csv("data/como_dispatch.csv", col_types = "cTcccccc")

# make it so that every column imports as a "character" column:
read_csv("data/como_dispatch.csv", col_types = cols(.default = "c"))

# here's another way to do that based off some internet research:
read_csv("data/como_dispatch.csv", col_types = cols(.default = col_character()))

# specify a data type for just one column (call it by name):
read_csv("data/como_dispatch.csv", col_types = cols(InNum = "c"))


```
