---
title: "R Notebook"
output: html_notebook
---

```{r}
# don't need to run this code again: 
install.packages("tidyverse").
```

We load libraries at the beginning of every script
```{r}
library(tidyverse)
```

Load some data
```{r}
weather <- read_csv("data/weather_events.csv")
```

Main tidyverse functions
```{r}
# SORTING
weather %>% arrange(desc(date))

weather %>% arrange(desc(injuries))

# FILTERING
weather %>% filter(type == "tornado") %>% arrange(desc(fatalities))

weather %>% filter(fatalities > 0)

# operators
# ==
# !=
# >, <, >=, <=
# AND, OR: &, |

weather %>% filter(type == "tornado" & fatalities > 0)

weather %>% filter(type == "tornado" | type == "wind")

# GROUPING
weather %>% 
  group_by(type) %>% 
  summarise(total_inj = sum(injuries)) %>% 
  arrange(desc(total_inj))

weather %>% 
  summarise(total_inj = sum(injuries)) %>% 
  arrange(desc(total_inj))

```

To answer these questions, you'll need the summary function that counts rows: n()
This function requires no arguments. ex: `summarise(number_of_events = n())`
Remember to show your work and answer the question (if you can). 

QUESTIONS FOR HOMEWORK: 
1. What was the most common weather event type in your home state?
2. How many injuries and fatalities were associated with it? 
3. Which weather event type has the highest fatality rate? 
4. Which county has the most weather events? (use `cnty1_fips`)
5. Are weather events increasing over time? 