---
title: "WNBA analysis"
output: 
---

Salary data: HerHoopStats (2024 salary, 2023 stats) Team rosters: ESPN Downloaded in January

```{r}
library(tidyverse)
```

Load our two tables:

```{r}
wnba_salaries <- read_csv("data/wnba_salaries.csv")
wnba_teams <- read_csv("data/wnba_teams.csv")
```

Practice joins

```{r}
anti_join(wnba_salaries, wnba_teams, by = c("name" = "player_name"))

# Isabelle Harrison & Lou Lopez Senechal not on team rosters (but have salaries?)

anti_join(wnba_teams, wnba_salaries, by = c("player_name" = "name"))

# AD Durr is on a team (Dream) but doesn't have a salary
```

Combine the tables:

```{r}
###LIZ: note that these two lines of code produce the same results: 

full_join(wnba_teams, wnba_salaries, by = c("player_name" = "name"))

wnba_teams %>% full_join(wnba_salaries, by = c("player_name"="name"))
```

Homework, due 2/19:

1.  Answer these questions
    1.  Which team has the highest average salary?
    2.  What team attempts the most field goals in a game?
    3.  Which team is the best rebounding team in the league?
    4.  What percentage of all players come from countries other than the US?
    5.  Which college produces the best players?
2.  Ask and answer three of your own questions.
